\chapter{Acknowledgments}
I extend my heartfelt gratitude to Professor Thomas Brox for his invaluable examination of this research thesis. His insightful feedback and guidance have played a pivotal role in shaping the outcome of this work. Additionally, I would like to express my sincere appreciation to Prof. Frank Hutter for sparking my interest in deep learning at the onset of my master's course and for examining my thesis.

I am deeply thankful for the unwavering support, mentorship, and expertise provided by my supervisors, Maria Bravo and Silvio Galesso, throughout the entire research process. Their continuous encouragement has been instrumental in driving my progress and enhancing the quality of this thesis.

The presence of my friends greatly enriched my research journey. Our collaborative sessions, shared experiences, and camaraderie played a pivotal role in shaping this meaningful academic endeavor. I want to extend my special thanks to Harsha for always being there with an open ear. Lastly, I am immensely grateful for my family's unwavering support; their belief in me and their love fueled my determination to achieve this significant academic milestone.
% First and foremost, I would like to thank...
% \begin{itemize}
% \item{advisers}
% \item{examiner}
% \item{person1 for the dataset}
% \item{person2 for the great suggestion}
% \item{proofreaders}
% \end{itemize}

%
% Feedback from Mria
% - add one qualitative example after every experiment
% - add little bit about cross attention, can reduce content about multihead or combine it together with self-attention as we are not changing anything related to it
%add label entropy loss say that insignificant performnace ensures the finding in motivation anyway
%add the same thingy about changing the hierarchy



%